# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

Reading data
```{r}
data <- read.csv(unz("activity.zip", "activity.csv"))
```

```{r}
head(data)
tail(data)
```

## What is mean total number of steps taken per day?

Average of each day
```{r}
average.day <- aggregate(steps~date, data=data, mean)
```

Total number of steps taken each day
```{r}
total.day <- aggregate(steps~date, data=data, sum)
```

Histogram of the total number of steps taken each day
```{r}
hist(total.day$steps)
```

Mean and median total number of steps taken per day
```{r}
mean(total.day$steps, na.rm=T)
median(total.day$steps, na.rm=T)
```

## What is the average daily activity pattern?

Time series plot of the 5-minute interval and the average number of steps taken
```{r}
average.interval <- aggregate(steps~interval, data=data, mean)
```

```{r}
plot(average.interval$steps, type="l")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
which.max(average.interval$steps)
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sapply(data, function(x) sum(is.na(x)))
```

For missing values I use *Multiple Imputation*.
Missing values are filled in based on non-missing values.
I use Multiple Imputation by Chained Equations (mice) package.

Install and load mice package
```{r}
library(mice)
```

Perform Multiple Imputation. Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
set.seed(144)
imputed = complete(mice(data))
```

Make a histogram of the total number of steps taken each day 
```{r}
steps.day <- aggregate(steps~date, data=imputed, sum)
```

```{r}
hist(steps.day$steps)
```

Calculate and report the mean and median total number of steps taken per day
```{r}
mean(steps.day$steps)
median(steps.day$steps)
```

The impact of imputing missing data on the estimates of the total daily number of steps is that the mean and median are slightly higher with imputation than without it

## Are there differences in activity patterns between weekdays and weekends?

Transform factor variable date in a *date format*  
```{r}
date.format <- as.Date(steps.day$date)
```

Getting names of the days
```{r}
days.name <- weekdays(date.format)
```

Index of weekends
```{r}
indx.weekend <- days.name == "Sunday" | days.name == "Saturday" 
```

Steps mean of weekends
```{r}
mean(steps.day$steps[indx.weekend]) 
```

Steps mean of no weekends days
```{r}
mean(steps.day$steps[-indx.weekend]) 
```

According to these results it seems the number of steps during weekends is higher.





